---
title: "Visualization"
subtitle: "Part 1"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data
So rnoaa is a publicly available source of data. The three titles in the first row c("shhg")are different weather stations on the public site and I'm getting info from them. We are looking at particular variables, precipitation and min and max temp. So it's getting this info from the publicly available site. The names are the locations of those weather stations. The Tmin was also divided by 10 because of how it was entered into the set.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```
## Scatterplots! 
Create my first scatterplot. You first enter the data set, then x and y. You add the geom point to get the scatterplot. 

```{r}
ggplot(weather_df, aes( x= tmin, y = tmax)) + 
  geom_point()
```
New approach, same plot
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + geom_point()
```
Save and edit a plot object: so just the plot without the data points

```{r}
weather_plot =
  weather_df %>% 
  ggplot(aes(x= tmin, y=tmax))
weather_plot
```
Now I can go back and add geom_point to that plot
```{r}
weather_plot + geom_point()
```
## Advanced scatterplots... 
Start with the same one and make it fancy! Now I'm telling it to assign a color based on names. 

```{r}
ggplot(weather_df, aes( x= tmin, y = tmax, color = name)) + 
  geom_point()
```
Now I'll add a lines that run through the middle of the data set for each color.He plotted with standard error being false, not quite sure what that means but you can still plot without that argument.  

```{r}
ggplot(weather_df, aes( x= tmin, y = tmax, color = name)) + 
  geom_point() + geom_smooth(se= FALSE)
```
What happens if I change the color aesthetic placement?
Here it creates a single line for all of the data because instead of assigning color to particular region, it puts it for the plot in general when it's in geom_point.

```{r}
ggplot(weather_df, aes(x= tmin, y = tmax)) + 
  geom_point(aes(color = name)) + geom_smooth(se= FALSE)
```
Let's facet some things! 

Going back to the original plot with separate smooth lines. Facet creates separate scatterplots for each name (. = don't create rows, the label after the ~ specifies that I want to separate into columns)
```{r}
ggplot(weather_df, aes( x= tmin, y = tmax, color = name)) + 
  geom_point() + geom_smooth(se= FALSE) +
  facet_grid(. ~ name)
```
If I change the order of the variables in the facet function, the display of the scatterplots will also vary. 

This time, I'm telling it to make 3 rows for my names, instead of columns for the names (name in the file corresponds to region). 

```{r}
ggplot(weather_df, aes( x= tmin, y = tmax, color = name)) + 
  geom_point() + geom_smooth(se= FALSE) +
  facet_grid(name ~.)
```
I can also make the individual data points themselves a bit more transparent so it's easier to see the line running through the middle of data. I change the `alpha` in the geom_point function or even adding `size` to the geom_point to change the size of the points
```{r}
ggplot(weather_df, aes( x= tmin, y = tmax, color = name)) + 
  geom_point(alpha =0.5, size = 0.4) + geom_smooth(se= FALSE) +
  facet_grid(. ~ name)
```
You can also change the transparency of the points based on the tmin, for example. So you see the gradient of temperatures
```{r}
ggplot(weather_df, aes( x= tmin, y = tmax, alpha =tmin, color = name)) + 
  geom_point( size = 0.4) + geom_smooth(se= FALSE) +
  facet_grid(. ~ name)
```
Let's combine some elements and try a new plot. 
Here, we plot the max temperatures (y) over dates (x). I also wanted to plot the size of the points based on the amount of precipitation at that time.And then of course if I wanted to create 3 separate plots based on the cities, I could use the facet function above. 

```{r}
weather_df %>% 
  ggplot(aes(x=date, y= tmax, color= name)) + geom_point( aes(size = prcp), alpha =0.5) +
geom_smooth(se = FALSE)
```
## Some small notes

There's also a hexagonal option for the shape of your points. You may need to install geom_hex if it asks, click 1 for yes. 

```{r}
weather_df %>% 
  ggplot(aes(x=date, y= tmax, color= name)) + geom_hex()
```
## Univariate Plots

Histograms

You can add color for each bar using this function

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill =name)) + geom_histogram()
```
However, they are overlapping, so I will tell the values to dodge each other

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill =name)) + geom_histogram(position = "dodge")

```
I could alternatively tell it to create separate plots
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill =name)) + geom_histogram() + facet_grid(. ~ name)
                                            
```
You can also make smoothed out histograms
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill =name)) + geom_density() 
                                            
```
Making Box plots

```{r}
weather_df %>% 
  ggplot(aes(x=name, y=tmin)) + geom_boxplot() 
```
Violin Plots 

You can add stat_summary to get the mean (by default if you don't set an argument). I have it set to find the median
```{r}
weather_df %>% 
  ggplot(aes(x=name, y=tmin, fill = name)) + geom_violin() + stat_summary( fun ="median")
```
Ridge plot
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=name)) + geom_density_ridges()
```

