---
title: "Visualization"
subtitle: "Part 2"
output: github_document
---

We load patchwork to make different types of plots
```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the weather data
So rnoaa is a publicly available source of data. The three titles in the first row c("shhg")are different weather stations on the public site and I'm getting info from them. We are looking at particular variables, precipitation and min and max temp. So it's getting this info from the publicly available site. The names are the locations of those weather stations. The Tmin was also divided by 10 because of how it was entered into the set.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```
## Remember this plot?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5)
  
```


## Labels

We can add labels to this plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
)
```
##Scales 

Start with the same plot, but now I can change the scale on each axis. I can also change the y axis so the values are plotted by the sqrt of the original. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
scale_x_continuous(
  breaks = c(-15, 0, 15),
  labels = c("-15 C", "0", "15")
) +
  scale_y_continuous(
    trans = "sqrt"
  )
```
 
 Adding color to scales. The numbers you enter into the hue function correspond to colors. The further the numbers are apart, I believe the further the difference in the hue. I am also telling the code to change name to location for the hues. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
  scale_color_hue(
    name = "Location",
    h= c(100, 200))
```
 You can also use colors from the viridis color scale- I guess the main benefit is for color blind patients, but truly I am not a fan of this scale. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

## Themes 

You can change the background of the plot, the location of legends, etc using the theme function. 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom") 
```

You can change the theme to black and white instead of gray in the background, and get rid of the box around the plot. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal()
```

BUT the order in which you place the code for theme matters. It will use the last code for theme and override anything else previously entered. Example, you'll see that the legend is not at the bottom anymore, though I've included it in the code. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom") +
  theme_minimal()
```

So you have to change the theme first
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha=0.5) +
labs(
  title = "Temperature plot",
  x = "Minimum daily temperature (C)",
  y = "Maximum daily temperature (C)",
  caption = "Data from the rnoaa package; temperatures in 2017"
) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom") 
  
```

## Setting options 

So this is a set of code you can run at the beginning of your file once you load in tinyverse so that the plots you make already have this organization. But we set the evaluation to false so it's not particularly run in this code. Asp = aspect ratio

```{r, eval= FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = 0.6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(ggplot2.continuous.color = "viridis", 
        ggplot2.continuous.fill ="viridis")

scale_color_discrete = scale_color_viridis_d()
scale_fill_discrete = scale_fill_viridis_d
```

## Data arguments in`geom`

I'm setting a data set for one set of data points, and then another data set to plot as a line. 

```{r}
central_park = 
  weather_df %>% 
  filter (name == "CentralPark_NY")
central_park

Molokai =
  weather_df %>% 
  filter (name == "Molokai_HI")
Molokai

ggplot(data = Molokai, aes(x = date, y = tmax, color = name))+ 
  geom_point() + geom_line(data = central_park)
  
```
## Learning about the `Patchwork` function

Recall the facet function separates plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) + 
  geom_density() +
facet_grid(.~name)
```
Sometimes you can't facet, so we can use patchwork for that instead. In this case, we entered code for 3 different plots, and made 1 a density plot with the other 2 as scatterplots. This code at the bottom `(tmax_tmin_p + prcp_dens_p) / tmax_date_p` tells R how I want these plots to be arranged-- two on top and then one on bottom. 

```{r}
tmax_tmin_p = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data manipulation 

GGplot automatically arranges variables in alphabetical order (as you can see in previous plots, central park is always first, then molokai then waterhole). You can use the factor function to change the categorization of name from a character variable to a factor variable, so that you can control how it's arranged. I told it to put Molokai_HI first. I will create a violin plot. For some reason, I used the same code

```{r}
weather_df %>% 
  mutate(
    name = as.factor(name),
    name = forcats::fct_relevel(name, c("Molokai_HI"))
  ) %>% 
    ggplot(aes(x= name, y=tmax, fill= name)) +
    geom_violin(alpha=0.5)
weather_df
```

